---
title: "carrerfour_analysis"
author: "Bill"
date: "11/13/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# 1. Problem Definition

## 1.1 Defining the Question

As a Data analyst at Carrefour Kenya,I have been consulted to undertake a project that will inform the marketing department on the most relevant marketing strategies that will result in the highest no. of sales (total price including tax). My  project will explore a recent marketing dataset provided by performing various unsupervised learning techniques and later providing recommendations based on my insights.

## 1.2 Specifying the Question
 I am expected  to check whether there are any anomalies in the given sales dataset. The objective of this task being fraud detection.

## 1.4 Understanding the Context

Carrefour, one of the largest hypermarket chains in the world was introduced to the Middle East and North Africa (MENA) market in 1995 by Majid Al Futtaim, the leading shopping mall, retail and leisure pioneer across MENA.

Carrefour has become the most dynamic, fast-moving and exciting hypermarket chain in the region and shared its growth with more than 38,000 employees from more than 70 nationalities in 15 countries, providing shoppers with variety and value-for-money. 

Carrefour ensures customer satisfaction and everyday convenience while offering unbeatable value for money with a vast array of more than 100,000 products, shoppers can purchase items for their every need, whether home electronics or fresh fruits from around the world, to locally produced items. 

Carrefour opened its first outlet in Kenya in 2016, and currently operates over 250 hypermarkets, supermarkets, and online stores in 15 countries across the region, with plans to extend into 38 countries in the Middle East, Central Asia, Africa and Russia.

Carrefour always strive to provide the best quality and most diverse selection of household goods available in Kenya. Our value packs and combination discount offers means that we can offer these products at even lower costs, keeping your household essentials at unbeatable prices.


## 1.5 Experimental Design taken
1. Data Exploration
2. Data Cleaning and Formatting
3. Univariate Analysis
4. Bivariate Analysis
5. Multivariate Analysis
6. Anomaly detection
6. Conclusion and Next steps
# 2. Data Sourcing
The data was availed to our data science team by the  Carrefour’s Sales and Marketing team therefore no data collection and scrapping was needed...We will just load our dataset in RStudio and begin the analysis process
# 3. Check the Data

``` {r}
# Load tidyverse and anomalize
# ---
# 
# load libraries
library(tidyverse)
library(tibbletime)
library(anomalize)
library(readr)
```


``` {r}
Supermarket_Sales_Forecasting_Sales <- read_csv("datasets/Supermarket_Sales_Forecasting - Sales.csv")
```

``` {r}
## previewing first 6 rows
head(Supermarket_Sales_Forecasting_Sales)

##previewing the last 6 rows of the dataset
tail(Supermarket_Sales_Forecasting_Sales)
```

``` {r}
## Previewing the shape of our dataset
dim(Supermarket_Sales_Forecasting_Sales)
### we have 1000  rows and 2 columns!!!BAM!!

#checking the datatypes on the columns 
sapply(Supermarket_Sales_Forecasting_Sales, class)

## the datecolumn is in character form,,, will need to convert it to datetime
Supermarket_Sales_Forecasting_Sales$Date <- anytime::anydate(Supermarket_Sales_Forecasting_Sales$Date)

str(Supermarket_Sales_Forecasting_Sales)

## We then a statistical summary of our dataset

summary(Supermarket_Sales_Forecasting_Sales)

```

Extract the year, month and day from the date column

```{r}
Supermarket_Sales_Forecasting_Sales$year <- as.numeric(format(Supermarket_Sales_Forecasting_Sales$Date, format="%Y"))
Supermarket_Sales_Forecasting_Sales$month <- as.numeric(format(Supermarket_Sales_Forecasting_Sales$Date, format="%m"))
Supermarket_Sales_Forecasting_Sales$day <- as.numeric(format(Supermarket_Sales_Forecasting_Sales$Date, format="%d"))
#check the datatypes
head(Supermarket_Sales_Forecasting_Sales)
```
## Data Cleaning

## To ensure uniformity, I will lowercase all the columns
``` {r}
names(Supermarket_Sales_Forecasting_Sales)<- tolower(names(Supermarket_Sales_Forecasting_Sales))
head(Supermarket_Sales_Forecasting_Sales) 


```

``` {r}
##Checking for missing values in each row
colSums(is.na(Supermarket_Sales_Forecasting_Sales))

```

``` {r}
anyDuplicated(Supermarket_Sales_Forecasting_Sales)

```

``` {r}
# Checking for Outliers in Sales column
boxplot(Supermarket_Sales_Forecasting_Sales) 
```
##EDA

``` {r}
## finding the summary statistic of our time series dataset
summary(Supermarket_Sales_Forecasting_Sales)



```

```{r}
# Correlations
correlations <- cor(Supermarket_Sales_Forecasting_Sales[,2:3], method = "pearson")
round(correlations, 2)  
```
# Confirms no relationship between day and sales

##Anomaly Detection

``` {r}

## Anomaly Detection
```{r}
# change the class of the data to tibble to be able to use the anomalize function
sales_anomaly <- as_tibble(Supermarket_Sales_Forecasting_Sales)
head(sales_anomaly)
```
```{r}
# check for anomalies in the Sales column using the anomalize() function
anomalies <- anomalize(sales_anomaly, sales, method = "iqr", alpha = 0.1, max_anoms = 0.2, verbose = FALSE)
head(anomalies)
```

```{r}
# check for unique values in the column "anomaly"
unique(anomalies$anomaly)
```

- Anomalies have been detected in our column anomaly.. This might be due to the presence of anomalies

```{r}
# convert the anomaly column to factor
data<- anomalies
data$anomaly <- as.factor(data$anomaly)
str(data)
```
## Plotting of anomalies

```{r}
#plot(sales_anomalies_data$Sales, color = "anomaly")
data %>% 
  ggplot() +
  aes(x = sales, anomaly = ..count../nrow(data), fill = anomaly) +
  geom_dotplot() +
  ylab("freq")
```

## Insights
_ There are anomalies in sales above 1000,,,,This should be checked by the company