---
title: "Association_analysis"
author: "Bill"
date: "11/13/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# 1. Problem Definition

## 1.1 Defining the Question

As a Data analyst at Carrefour Kenya,I have been consulted to undertake a project that will inform the marketing department on the most relevant marketing strategies that will result in the highest no. of sales (total price including tax). My  project will explore a recent marketing dataset provided by performing various unsupervised learning techniques and later providing recommendations based on my insights.

## 1.2 Specifying the Question
 I am expected to  identify relationships between variables in the dataset and  to provide insights for my analysis.

## 1.4 Understanding the Context

Carrefour, one of the largest hypermarket chains in the world was introduced to the Middle East and North Africa (MENA) market in 1995 by Majid Al Futtaim, the leading shopping mall, retail and leisure pioneer across MENA.

Carrefour has become the most dynamic, fast-moving and exciting hypermarket chain in the region and shared its growth with more than 38,000 employees from more than 70 nationalities in 15 countries, providing shoppers with variety and value-for-money. 

Carrefour ensures customer satisfaction and everyday convenience while offering unbeatable value for money with a vast array of more than 100,000 products, shoppers can purchase items for their every need, whether home electronics or fresh fruits from around the world, to locally produced items. 

Carrefour opened its first outlet in Kenya in 2016, and currently operates over 250 hypermarkets, supermarkets, and online stores in 15 countries across the region, with plans to extend into 38 countries in the Middle East, Central Asia, Africa and Russia.

Carrefour always strive to provide the best quality and most diverse selection of household goods available in Kenya. Our value packs and combination discount offers means that we can offer these products at even lower costs, keeping your household essentials at unbeatable prices.


## 1.5 Experimental Design taken
1. Data Exploration
2. Data Cleaning and Formatting
3. Univariate Analysis
4. Bivariate Analysis
5. Multivariate Analysis
6. Associative Analysis
6. Conclusion and Next steps
# 2. Data Sourcing
The data was availed to our data science team by the  Carrefour’s Sales and Marketing team therefore no data collection and scrapping was needed...We will just load our dataset in RStudio and begin the analysis process
# 3. Check the Data

``` {r}
## Loading packages that we will use during our analysis
library("dplyr")
library("purrr")
library('tidyverse')
library('magrittr')
library('corrplot')
library('caret')
library('skimr')
library(readr)
library(arules)

```
``` {r}
path <-"datasets/Supermarket_Sales_Dataset II.csv"

supermarket<-read.transactions(path, sep = ",")

supermarket

```
- Our dataset has 7501 transactions  and 119 items

``` {r}
# Verifying the object's class
# ---
# This should show us transactions as the type of data that we will need
# ---
# 
class(supermarket)



```

``` {r}
# Previewing our first 5 transactions
#
inspect(supermarket[1:5])

```

``` {R}
# Generating a summary of the transaction dataset
# ---
# This would give us some information such as the most purchased items, 
# distribution of the item sets (no. of items purchased in each transaction), etc.
# ---
# 
summary(supermarket)
```


## Observations
1. Mineral water was the most purchased item with a total number of 1788
2. eggs were the next purchased item in the supermarket with a total number of 1348   
3. Sphagetti, french fries and chocolate were the next most purchased items respectively


``` {r}
# Exploring the frequency of some items in our supermarket 
# i.e. goods ranging from 8 to 10 and performing 
# some operation in percentage terms of the total transactions 
# 
itemFrequency(supermarket[, 8:10],type = "absolute")

round(itemFrequency(supermarket[, 8:10],type = "relative")*100,2)
```



``` {r}
# Producing a chart of frequencies and fitering 
# to consider only items with a minimum percentage 
# of support/ considering a top x of items
# ---
# Displaying top 10 most common items in the transactions dataset 
# and the items whose relative importance is at least 10%
# 
par(mfrow = c(1, 2))

# plot the frequency of items
itemFrequencyPlot(supermarket, topN = 20,col="darkgreen")
itemFrequencyPlot(supermarket, support = 0.1,col="darkred")
```
``` {r}

itemFrequencyPlot(supermarket, support = 0.1,col="darkred")
```


``` {r}
# Building a model based on association rules 
# using the apriori function 
# ---
# We use Min Support as 0.001 and confidence as 0.8
# ---
# 
rules <- apriori (supermarket, parameter = list(supp = 0.001, conf = 0.8))
rules

```

``` {r}
# We can perform an exploration of our model 
# through the use of the summary function as shown
# ---
# Upon running the code, the function would give us information about the model 
# i.e. the size of rules, depending on the items that contain these rules. 
# More statistical information such as support, lift and confidence is also provided.
# ---
# 
summary(rules)
```


``` {r}
# Observing rules built in our model i.e. first 15 model rules
# ---
# 
inspect(rules[1:15])



```
##Observations
-if someone buys  frozen smoothie,spinach, there is an 88% confidence that he will buy mineral water
- if someone buys bacon,pancakes, there is an 81% confidence that he will buy spaghetti
- if someone buys mushroom cream sauce,pasta, there is an 95% confidence that he will buy escalope
``` {r}
# Ordering these rules by a criteria such as the level of confidence
# then looking at the first five rules.
# We can also use different criteria such as: (by = "lift" or by = "support")
# 
rules<-sort(rules, by="confidence", decreasing=TRUE)
inspect(rules[1:15])

# Interpretation
# ---
# The given five rules have a confidence of 100
# ---
```
```  {r}
library(arulesViz)
plot(rules[1:15],
method = "graph",
control = list(type = "items"))
```



``` {r}
# If we're interested in making a promotion relating to the sale of mineral water, 
# we could create a subset of rules concerning these products 
# ---
# This would tell us the items that the customers bought before purchasing mineral water
# ---
# 
mineral_water <- subset(rules, subset = rhs %pin% "mineral water")
 
# Then order by confidence
mineral_water<-sort(mineral_water, by="confidence", decreasing=TRUE)
inspect(mineral_water[1:5])
```

## Observations
- ground beef, light cream, olive oil were bought before mineral water was purchased
- cake, olive oil, shrimp were also bought before mineral water was purchased
- red wine and soup were also bought before mineral water was purchased
- frozen vegetables, milk, spaghetti,turkey were also bought before mineral water was purchased

``` {r}
# If we're interested in making a promotion relating to the sale of milk, 
# we could create a subset of rules concerning these products 
# ---
# This would tell us the items that the customers bought before purchasing eggs 
# ---
# 
milk <- subset(rules, subset = rhs %pin% "milk")
 
# Then order by confidence
mlk<-sort(milk, by="confidence", decreasing=TRUE)
inspect(milk[1:5])
```
## Observations
- cake,meatballs,mineral water, escalope,hot dogs, meatballs,whole wheat pasta, black tea,frozen smoothie and burgers,ground beef,olive oil were bought before milk


## Insights
1. Mineral water was the most purchased item with a total number of 1788
2. eggs were the next purchased item in the supermarket with a total number of 1348   
3. Sphagetti, french fries and chocolate were the next most purchased items respectively
4. Most people buy 3-5 items per transaction with majority being 4 items per transaction as per the rule length distribution
5.  The following items should be placed near  each other
- frozen smoothie,spinach
- red wine,tomato sauce
- black tea,frozen smoothie
- black tea,salmon
- cooking oil,fromage blanc
- milk,pasta
- mushroom cream sauce,pasta
- ground beef,nonfat milk
- bacon,pancakes
- nonfat milk,turkey
- turkey,whole wheat pasta
- red wine,soup
- cake,meatballs,mineral water
- chicken,protein bar
- meatballs,whole wheat pasta